### 2014-07-13 doc/manual/overview.tex
Initial commit

\chapter{Overview}

The prevalent uses of JavaScript in web programming have revealed
security vulnerability issues of JavaScript applications,
which emphasizes the need for JavaScript analyzers to detect such issues.
Recently, researchers have proposed several analyzers of JavaScript programs
and some web service companies have developed various JavaScript engines.
However, unfortunately, most of the tools are not documented well,
thus it is very hard to understand and modify them.
Or, such tools are often not open to the public.

In this specification, we present formal specification and implementation
of \safe, a scalable analysis framework for ECMAScript, developed for the JavaScript research community.
This is the very first attempt to provide both formal specification and
its open-source implementation for JavaScript,
compared to the existing approaches focused on only one of them.
To make it more amenable for other researchers to use our framework,
we formally define three kinds of intermediate representations for JavaScript
used in the framework,
and we provide formal specifications of translations between them.
To be adaptable for adventurous future research including modifications
in the original JavaScript syntax,
we actively use open-source tools to automatically generate parsers and some intermediate
representations.
To support a variety of program analyses in various compilation phases,
we design the framework to be as flexible, scalable, and pluggable as possible.
Finally, our framework is publicly available,
and some collaborative research using the framework are in progress.