### 2016-05-05 doc/overview.tex
Add the SAFE document till AST2IR

\chapter{Overview}

The prevalent uses of JavaScript in web programming have revealed
security vulnerability issues of JavaScript applications,
which emphasizes the need for JavaScript analyzers to detect such issues.
Recently, researchers have proposed several analyzers of JavaScript programs
and some web service companies have developed various JavaScript engines.
However, unfortunately, most of the tools are not documented well,
thus it is very hard to understand and modify them.
Or, such tools are often not open to the public.

In this specification, we present formal specifications and implementation
of \safe, a scalable analysis framework for ECMAScript, developed for the JavaScript research community.
To make it more amenable for other researchers to use our framework,
we formally define three kinds of intermediate representations for JavaScript
used in the framework,
and we provide formal specifications of translation between them.
To support a variety of program analyses in various compilation phases,
we design the framework to be as flexible, scalable, and pluggable as possible.
Finally, our framework is publicly available:
\begin{center}
\url{http://safe.kaist.ac.kr}
\end{center}
Our project has been supported by
Korea Ministry of Education, Science and Technology (MEST),
National Research Foundation of Korea (NRF),
Samsung Electronics,
S-Core,
Microsoft Research Asia, and
Google.

Projects and colleagues using \safe\ include:
\begin{itemize}
\item K Framework @ University of Illinois at Urbana-Champaign
\item Ken Cheung @ HKUST
\item ROSAEC @ Seoul National University
\item Tizen @ Linux Foundation
\item Web-based Vulnerability Detection @ Oracle Labs
\end{itemize}
Publications related to \safe\ include:
\begin{itemize}
\item Battles with false positives in static analysis of JavaScript web applications in the wild (ICSE SEIP 2016)
\item Static analysis of JavaScript web applications in the wild via practical DOM modeling (ASE 2015)
\item Practically tunable static analysis framework for large-scale JavaScript applications (ASE 2015)
\item Development nature matters: An empirical study of code clones in JavaScript applications (EMSE 2015)
\item Scalable and precise static analysis of JavaScript applications via loop-sensitivity (ECOOP 2015)
\item {\sf SAFE$_{\mbox{\sf\scriptsize WAPI}}$}: Web API misuse detector for web applications (FSE 2014)
\item All about the \texttt{with} statement in JavaScript: Removing \texttt{with} statements in JavaScript applications (DLS 2013)
\item Formal specification of a JavaScript module system (OOPSLA 2012)
\item \safe: Formal specification and implementation of a scalable analysis framework for ECMAScript (FOOL 2012)
\end{itemize}