### 2016-10-28 doc/safe-ir.tex

\chapter{IR}
This section describes each construct of the \safe IR
in both the BNF notation and its corresponding implementation.
The implementation of IR nodes is available at:
\begin{verbatim}
$SAFE_HOME/src/main/scala/kr/ac/kaist/safe/nodes/ir/
\end{verbatim}
\small
\[
\begin{array}{llll}
\ir\pgm & ::= & \ir\stmt^* \\
&& \mtt{IRRoot(val fds:~List[IRFunDecl],}\\
&&\mtt{\phantom{IRRoot(}val vds:~List[IRVarStmt],}\\
&&\mtt{\phantom{IRRoot(}val irs:~List[IRStmt])}\\\\

\ir\stmt & ::= & \irid \ \verb+=+ \ \ir\expr \\
&& \mtt{IRExprStmt(val lhs:~IRId, val right:~IRExpr,}\\
&&\mtt{\phantom{IRExprStmt(}val ref:~Boolean)}\\
&\mid& \irid \ \verb+=+ \ {\sf delete}\ \irid\\
&& \mtt{IRDelete(val lhs:~IRId, val id:~IRId)}\\
&\mid& \irid \ \verb+=+ \ {\sf delete}\ \irid\verb+[+\irid\verb+]+\\
&& \mtt{IRDeleteProp(val lhs:~IRId, val obj:~IRId,}\\
&&\mtt{\phantom{IRDeleteProp(}val index:~IRExpr)}\\
&\mid& \irid\verb+[+\irid\verb+] =+ \ \irexpr \\
&& \mtt{IRStore(val obj:~IRId, val index:~IRExpr,}\\
&&\mtt{\phantom{IRStore(}val rhs:~IRExpr)}\\
 &\mid& \irid \ \verb+=+ \ \verb+{+ (\ir\member\verb+,+)^* \verb+}+\\
&& \mtt{IRObject(val lhs:~IRId, val members:~List[IRMember],}\\
&&\mtt{\phantom{IRObject(}val proto:~Option[IRId])}\\
 &\mid& \irid \ \verb+=+ \ \verb+[+ (\irexpr\verb+,+)^* \verb+]+ \\
&& \mtt{IRArray(val lhs:~IRId,}\\
&&\mtt{\phantom{IRArray(}val elements:~List[Option[IRExpr]])}\\
&&\mtt{IRArgs(val lhs:~IRId,}\\
&&\mtt{\phantom{IRArray(}val elements:~List[Option[IRExpr]])}\\
 &\mid& \irid \ \verb+=+ \ \irid\verb+(+\irid\verb+,+\irid\verb+)+\\
&& \mtt{IRCall(val lhs:~IRId, val fun:~IRId,}\\
&&\mtt{\phantom{IRCall(}val thisB:~IRId, val args:~IRId)}\\
 &\mid& \irid \ \verb+=+ \ \irid\verb+(+\irid(\verb+,+\irid)^?\verb+)+\\
&& \mtt{IRInternalCall(val lhs:~IRId, val fun:~IRId,}\\
&&\mtt{\phantom{IRInternalCall(}val first:~IRExpr,}\\
&&\mtt{\phantom{IRInternalCall(}val second:~Option[IRId])}\\
&&{\inblue \mtt{toObject}, \mtt{toNumber}, \mtt{isObject},
\mtt{getBase}, \mtt{iteratorInit},}\\
&&{\inblue \mtt{iteratorHasNext}, \mtt{iteratorKey}}\\

 &\mid& \irid \ \verb+=+ \ {\sf new}\ \irid\verb+(+(\irid\verb+,+)^*\verb+)+\\
&& \mtt{IRNew(val lhs:~IRId, val fun:~IRId,}\\
&&\mtt{\phantom{IRNew(}val args:~List[IRId])}\\
\end{array}
\]

\[
\begin{array}{llll}
 &\mid& \irid \ \verb+=+ \ {\sf function} \ \ir{f} \verb+(+\irid\verb+,+\irid\verb+) {+ \ir\stmt^* \verb+}+\\
&& \mtt{IRFunExpr(val lhs:~IRId, val fun:~IRFunctional)}\\
&&\mtt{IRFunctional(val fromSource:~Boolean,}\\
&&\mtt{\phantom{IRFunctional(}val name:~IRId,}\\
&&\mtt{\phantom{IRFunctional(}val params:~List[IRId],}\\
&&\mtt{\phantom{IRFunctional(}val args:~List[IRStmt],}\\
&&\mtt{\phantom{IRFunctional(}val fds:~List[IRFunDecl],}\\
&&\mtt{\phantom{IRFunctional(}val vds:~List[IRVarStmt],}\\
&&\mtt{\phantom{IRFunctional(}val body:~List[IRStmt])}\\
 &\mid& {\sf function} \ \ir{f} \verb+(+\irid\verb+,+\irid\verb+) {+ \ir\stmt^* \verb+}+\\
&& \mtt{IRFunDecl(val ftn:~IRFunctional)}\\
 &\mid& \irid \ \verb+=+ \ {\sf eval}\verb+(+\ir\expr\verb+)+ \\
&& \mtt{IREval(val lhs:~IRId, val arg:~IRExpr)}\\
 &\mid& {\sf break} \ \irid \\
&& \mtt{IRBreak(val label:~IRId)}\\
 &\mid& {\sf return} \ \irexpr^?\\
&& \mtt{IRReturn(val expr:~Option[IRExpr])}\\
 &\mid& {\sf with} \ \verb+(+\irid\verb+)+ \ \ir\stmt\\
&& \mtt{IRWith(val id:~IRId, val stmt:~IRStmt)}\\
 & \mid & \ir{l} \; \verb+: {+ \; \ir\stmt \; \verb+}+\\
 && \mtt{IRLabelStmt(val label:~IRId, val stmt:~IRStmt)}\\
 &\mid& {\sf var} \ \irid\\
&& \mtt{IRVarStmt(val lhs:~IRId, val fromParam:~Boolean)}\\
 &\mid& {\sf throw} \ \irexpr\\
&& \mtt{IRThrow(val expr:~IRExpr)}\\
 &\mid& {\ir\stmt}^*\\
&& \mtt{IRSeq(val stmts:~List[IRStmt])}\\
 &\mid& {\sf if} \ \verb+(+\irexpr\verb+)+\ {\sf then} \ \ir\stmt \ ({\sf else} \ \ir\stmt)^?\\
&& \mtt{IRIf(val expr:~IRExpr, val trueB:~IRStmt,}\\
&&\mtt{\phantom{IRIf(}val falseb:~Option[IRStmt])}\\
 &\mid& {\sf while} \ \verb+(+\irexpr\verb+)+\ \ir\stmt\\
&& \mtt{IRWhile(val cond:~IRExpr, val body:~IRStmt)}\\
%
 &\mid& {\sf try} \ \verb+{+ \ir\stmt \verb+}+ \
({\sf catch} \ \verb+(+\irid\verb+){+ \ir\stmt \verb+}+)^? ({\sf finally} \ \verb+{+ \ir\stmt \verb+}+)^?\\
&& \mtt{IRTry(val body:~IRStmt, val name:~Option[IRId],}\\
&&\mtt{\phantom{IRTry(}val catchB:~Option[IRStmt],}\\
&&\mtt{\phantom{IRTry(}finallyB:~Option[IRStmt])}\\
&\mid& \open {\ir\stmt}^* \close\\
&& \mtt{IRStmtUnit(List[IRStmt] stmts)}\\\\
\ir\expr &::=&
 \irexpr \ \inop \irexpr \\
&& \mtt{IRBin(val first:~IRExpr, val op:~IROp,}\\
&& \mtt{\phantom{IRBin(}val second:~IRExpr)}\\
 &\mid& \preop \irexpr \\
&& \mtt{IRUn(val op:~IROp, val expr:~IRExpr)}\\
 &\mid& \irid\verb+[+\ir\expr\verb+]+ \\
&& \mtt{IRLoad(val obj:~IRId, val index:~IRExpr)}\\
 &\mid& \irid\\
&& \mtt{IRUserId(val global:~Boolean, val isWith:~Boolean)}\\
 &\mid& \newvar{x}\\
&& \mtt{IRTmpId(val global:~Boolean)}\\
 &\mid& \ir\num \\
&& \mtt{IRNumber(val text:~String, val num:~Double)}\\
 &\mid& \ir\str \\
&& \mtt{IRString(val str:~String)}\\
 &\mid& {\sf true} \\
&& \mtt{IRBool(val bool:~Boolean)}\\
 &\mid& {\sf false} \\
&& \mtt{IRBool(val bool:~Boolean)}\\
\end{array}
\]

\[
\begin{array}{llll}
 &\mid& {\sf undefined} \\
&& \mtt{IRUndef()}\\
 &\mid& {\sf null} \\
&& \mtt{IRNull()}\\
 &\mid& {\sf this} \\
&& \mtt{IRThis()}\\\\

\ir\member &::=& \irid \ \verb+:+ \ \irexpr\\
&& \mtt{IRField(val prop:~IRId, val expr:~IRExpr)}\\
 &\mid& {\tt get}\ \ir{f} \verb+(+\irid\verb+,+\irid\verb+) {+ \ir\stmt^* \verb+}+\\
&&\mtt{IRGetProp(val ftn:~IRFunctional)}\\
 &\mid& {\tt set}\ \ir{f} \verb+(+\irid\verb+,+\irid\verb+) {+ \ir\stmt^* \verb+}+\\
&&\mtt{IRSetProp(val ftn:~IRFunctional)}\\
\end{array}
\]

The \safe IR has the following assumptions and notations:
\begin{itemize}
\item We denote a list as a possibly empty, semicolon-separated sequence, enclosed by $\langle$ and $\rangle$.
\item We denote a series of list appends as superscripted $*$ such as $\stmt^*$.
\item We abuse our notations by interchanging semicolon-separated sequences and lists.
\item To denote an AST-level statement granularity in the translated IR statements,
we use {\tt IRStmtUnit} which is represented as green angle brackets {\ingreen $\open\ \close$} in this document.
\item Declarations of functions and variables are hoisted to their closest enclosing functions
or the top level via the \verb!Hoister! transformation of the \verb!astRewriter! phase.
\item Identifiers and labels that exist in the source program,
except when they appear at top level or within the {\tt with} statement,
are already disambiguated via the \verb!Disambiguator! transformation of the \verb!astRewriter! phase
so that they have unique names.
\end{itemize}