### 2020-09-01 paper/pldi21/abstract.tex
Revised the abstract

\begin{abstract}
Static analyses based on abstract interpretation can be integrated into developer workflow and help developers to find bugs.
However, the dynamic features of JavaScript degrade the scalability and precision of the pure static analyses.
Combining static and dynamic analyses is one of the breakthroughs in JavaScript bug detection.
Existing combined analyses have static boundaries that distinguish which analysis used for which part of a program.
Such boundaries are intuitive but limit chances to optimize transitions between two analyses.
We formally present the alloyed analysis, a more flexible compound of static and dynamic analyses.
We implement a proof of concept bug detector by extending the existing static analyzer supporting a subset of ECMAScript 5.1.
Our experimental results show that the alloyed analysis indeed speeds up the detection and reports bugs more precisely.
\end{abstract}