### 2020-09-09 paper/pldi21/detail.tex
[PLDI21] Removed unnecessary files / Updated formalization

\section{Application: JavaScript}

\todo

% \subsection{Core Language}
% \[
%   \begin{array}{ll@{~}c@{~}l}
%     \text{Programs} & \prog &::=& (\lab: \inst)^*\\
%     \text{Labels} & \lab &\in& \labset\\
%     \text{Instructions} & \inst &::=& x = \expr \mid \kwif \; \expr \; \lab \;
%     \lab\\
%     \text{Expressions} & \expr &::=& \const \mid x \mid \op(\expr, \cdots,
%     \expr)\\
%   \end{array}
% \]
% 
% In a core language, we define a program as a sequence of labelled instructions
% and instructions are variable assignments and branches.  An expression is a
% constant $\const$, a variable $x$, and $n$-ary operations.  We define states
% $\stset = \labset \times \memset$ as pairs of labels and memories mapping from
% variables to values.  We define the transition relation $\trans$ between states
% as follows:
% 
% \begin{prooftree}
%   \AxiomC{$\prog = \cdots, \lab: x = \expr, \lab': \inst, \cdots$}
%   \AxiomC{$\sem{e}(\mem) = v$}
%   \BinaryInfC{$(\lab, \mem) \trans (\lab', \mem[x \mapsto v])$}
% \end{prooftree}