### 2020-09-09 paper/pldi21/related.tex

\section{Related Work}

\todo

\subsection{Combined Analysis}

\begin{itemize}
  \item Concerto: a framework for combined concrete and abstract
    interpretation~\cite{concerto}
  \item Battles with false positives in static analysis of JavaScript web
    applications in the wild~\cite{battles}
  \item EventHandler-Based Analysis Framework for Web Apps Using Dynamically
    Collected States~\cite{eha}
  \item Automatic Modeling of Opaque Code for JavaScript Static
    Analysis~\cite{opaque-model}
  \item Mimic: Computing models for opaque code~\cite{mimic}
\end{itemize}

\subsection{Abstract Counting}

\begin{itemize}
  \item Improving flow analyses via $\Gamma$CFA: Abstract garbage collection and
    counting~\cite{abstract-gc-counting}
  \item Revisiting recency abstraction for JavaScript: towards an intuitive,
    compositional, and efficient heap abstraction~\cite{revisit-recency}
\end{itemize}

\subsection{Access Analysis}

\begin{itemize}
  \item Access Analysis-Based Tight Localization of Abstract
    Memories~\cite{func-local}
  \item Design and implementation of sparse global analyses for C-like
    languages~\cite{sparse}
\end{itemize}