### 2020-10-26 paper/pldi21/motivation.tex
Motivating example - Figure 1

\section{Motivation}\label{sec:motivation}

\begin{figure}[t]
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \begin{lstlisting}[style=myJSstyle]
function concat() {
  var length = arguments.length;
  if (!length) return [];
  var args = Array(length - 1),
      array = arguments[0],
      index = length;
  while (index--)
    args[index-1] = arguments[index];
  return arrayPush(isArray(array) ?
    copyArray(array) :
    [array], baseFlatten(args, 1));
}
    \end{lstlisting}
    \label{fig:example-concat}
    \vspace*{-1em}
    \caption{Lodash's \code{concat} function.}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\textwidth}
    \begin{lstlisting}[style=myJSstyle,firstnumber=13]
function getData(e) {
  var option = ... // option of server connection
  post(option).then(function(e) {
    if (e.total_records && e.total_records > 0) {
      // non-deterministic arguments of `concat`
      this.pastEvents =
        _.concat(this.pastEvents, e.events);
      this.total = e.total_records;
    } else this.noPastData = !0
  })
}
    \end{lstlisting}
    \label{fig:example-concat}
    \vspace*{-1em}
    \caption{The \code{getData} function to load more events from the server.}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\textwidth}
    \begin{lstlisting}[style=myJSstyle,firstnumber=24]
function changeCountry(G) {
  ...
  if (G.selectedVal === "US" && state) {
    // deterministic arguments of `concat`
    state.items = _.concat([["Other", "Other"]],
      WebinarBase.questions.state.items);
    state.selectedVal = _.head(_.head(C.items));
  }
}
    \end{lstlisting}
    \label{fig:example-concat}
    \vspace*{-1em}
    \caption{The \code{changeCountry} function to update state information.}
  \end{subfigure}
  \label{fig:example}
  \vspace*{-1em}
  \caption{Motivating Example: Excerpts from Lodash library and JavaScript codes
  in \code{zoom.us} site.}
\end{figure}

\begin{lstlisting}[style=myJSstyle]
WebinarBase.questions = { ...
  state: { ...
    items: [["AL","Alabama"], ["AK","Alaska"], ...]
  }
};
\end{lstlisting}

\todo