### 2020-10-29 paper/pldi21/motivation.tex
Revising motivation.

\section{Motivation}\label{sec:motivation}

\begin{figure}[t]
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \begin{lstlisting}[style=myJSstyle]
function concat() {
  var length = arguments.length;
  if (!length) return [];
  var args = Array(length - 1),
      array = arguments[0],
      index = length;
  while (index--)
    args[index-1] = arguments[index];
  return arrayPush(
    isArray(array) ? copyArray(array) : [array],
    baseFlatten(args, 1));
}
    \end{lstlisting}
    \vspace*{-1em}
    \caption{Lodash's \jscode{concat} function.}
    \label{fig:concat}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\textwidth}
    \begin{lstlisting}[style=myJSstyle,firstnumber=13]
function changeCountry(G) {
  ...
  if (G.selectedVal === "US" && state) {
    // deterministic arguments of `concat`
    state.items = _.concat([["Other", "Other"]],
      WebinarBase.questions.state.items);
    state.selectedVal = _.head(_.head(C.items));
  }
}
    \end{lstlisting}
    \vspace*{-1em}
    \caption{Load the list of states of the United States.}
    \label{fig:changeCountry}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\textwidth}
    \begin{lstlisting}[style=myJSstyle,firstnumber=22]
function getData(e) {
  var option = ... // option of server connection
  post(option).then(function(e) {
    if (e.total_records && e.total_records > 0) {
      // non-deterministic arguments of `concat`
      this.pastEvents =
        _.concat(this.pastEvents, e.events);
      this.total = e.total_records;
    } else this.noPastData = !0
  })
}
    \end{lstlisting}
    \vspace*{-1em}
    \caption{Load more Zoom events from the server.}
    \label{fig:getData}
  \end{subfigure}
  \vspace*{-1em}
  \caption{Motivating Example: Excerpts from Lodash library and JavaScript codes
  in \code{zoom.us} site.}
  \label{fig:example}
\end{figure}

In this section, we first point out limitations that existing techniques face to
analyze the real-world example described in Figure~\ref{fig:example}.
We then show how our combined analysis increase the analysis scalability and
precision through sealed symbolic execution, seamlessly combined with the original
static analysis.

We excerpt the \jscode{concat} function in Figure~\ref{fig:concat} from Lodash
library~\cite{lodash} (v4.17.20), which is the most popular npm package~\footnote{https://www.npmjs.com/browse/depended}
that \inred{124,562} npm packages depend on.
The \jscode{concat} function creates a new array concatenating given arrays or
values.
It first checks the length of arguments in line 2-3. Then, it stores the first
argument to \jscode{array} in line 5 and copies the remaining arguments to
\jscode{args} in line 7-8.
Finally, it creates a new array by copying the given array via \jscode{copyArray}
or initializing with a single value in line 10, and pushes each element of
\jscode{args} after flattening it via \jscode{baseFlatten} in line 11.

The functions in Lodash are popular to implement common functionalities for
websites as well.
We excerpt two functions \jscode{changeCountry} in Figure~\ref{fig:changeCountry}
and \jscode{getData} in Figure~\ref{fig:getData} from the \code{zoom.us}~\cite{zoom} site.
The website \code{zoom.us} is ranked at \inred{16th} popular web site accroding
to Alexa~\footnote{https://www.alexa.com/siteinfo/zoom.us} in November 2020.
The \jscode{changeCountry} is the example of using \jscode{concat} only with
concrete values.
The \jscode{changeCountry} function is invoked when a user selects his/her
country on the dropdown list in the register page.
When the user selects ``United States of America'', \jscode{changeCountry} calls
the \jscode{concat} function in line 17-18 to prepare the array for
``State/Province'' dropdown list.
The first argument of \jscode{concat} is an array literal \jscode{[["Other",
"Other"]]} and the second one is an array of pairs of abbreviations and names of
the states defined as follows:
\begin{lstlisting}[style=myJSstyle,numbers=none]
WebinarBase.questions.state.items =
  [["AL","Alabama"], ..., ["WY", "Wyoming"]]
\end{lstlisting}
Thus, \jscode{changeCountry} always invokes \jscode{concat} with concrete values.
In the case of \jscode{getData}, \jscode{concat} works with unknown values.
The \jscode{getData} function is invoked when loading more Zoom events in
``Webinars \& Events'' page.
At the first time, initial eight events appear on the page.
For each click of the ``Load More'' button, the \jscode{getData} sends a POST
request to the server and receives data of 8 more events by calling \jscode{concat} in line 28.
As a result, the events list on the page keeps growing by 8.
Both functions, selecting an option of dropdown list and loading more contents
from server-side, are familiar to users.

Existing techniques: static analysis, dynamic analysis, and combined analysis
have limitations to deal with the program containing \jscode{changeCountry} and
\jscode{getData}.
Static analyses of low degrees of context sensitivity tend to cause the
precision loss during analyzing \jscode{concat} function.
The loop in line 7-8 can lead to messing up every property value of args object.
Moreover, \jscode{baseFlatten} has a recursive call that is another challenge
for static analysis.
High degrees of context sensitivity is the promising way, but its computaion
cost is the burden.
Dynamic analyses can show much higher performance for the \jscode{concat} call
by \jscode{changeCountry} than static analyses without any unsoundness and
precision loss.
However, they cannot cover all possible behavior of the \jscode{concat} call
by \jscode{getData} originated by unknown server-side data.
A promising form of combined analysis is dynamically analyzing the call by \jscode{changeCountry}
and statically analyzing the call by \jscode{getData} that is not possible by
existing combined analyses.
User events trigger both functions, the order of function calls is
non-deterministic, so the proper point to divide stages for combined analyses
does not exist.
The syntactic division of the program imposes to choose only one kind of analysis
for the \jscode{concat} function that restricts the benefits of combined analysis.

We present a more flexible and fine-grained method to combine two analyses to
maximize the benefits.

Our observation is

%When the given arguments of a function invokation are concrete values, we can
%perform dynamic analysis instead of static analysis. For example, The
%\jscode{changeCountry} function is invoked when a country different with the
%current one is selected in registration of Zoom meetings.  When the ``United
%States of America'' (USA) is selected, it calls the \jscode{concat} function
%with two pre-defined concrete values as arguments to load USA state information
%in line 17-18.  The first argument is an array literal \jscode{[["Other",
%"Other"]]} and the second one is an array of pairs of abbreviations and names of
%USA states defined as follows:
%\begin{lstlisting}[style=myJSstyle,numbers=none]
%WebinarBase.questions.state.items =
%  [["AL","Alabama"], ..., ["WY", "Wyoming"]]
%\end{lstlisting}
%Moreover, \jscode{this} value is also a concrete value, the Lodash top-level
%object \jscode{\_}.  Thus, we could perform dynamic analysis by invoking the
%\jscode{concat} funciton with \jscode{\_} as \jscode{this} value and above two
%concrete values as arguments.  It increases performance of static analysis by
%skipping the analysis of function call in line 17-18 and utilizing the result of
%dynamic analysis.
%
%Dynamic analysis is still applicable using lazy concrete execution even if the
%arguments are not concrete values.  The \jscode{getData} function is another
%part to use \jscode{concat} function but we can only partially apply dynamic
%analysis in this case.  It is invoked when loading more Zoom events in
%``Webinars \& Events'' page.  At the first time, initial eight events are stored
%in \jscode{this.pastEvents}.  For each click of the ``Load More'' button, the
%\jscode{getData} sends a POST request to the server and receives additional
%event information \jscode{e} in line 24.  Then, eight events in
%\jscode{e.events} are appended to \jscode{this.pastEvents} using the
%\jscode{concat} function in line 27-28.  In this case, arguments are not
%deterministic but dependent on the given data from the server.  However, we know
%that the number of arguments are still deterministically defined as \jscode{2}
%and they are array objects.  Thus, we just invoke the funciton \jscode{concat}
%with \jscode{\_} as \jscode{this} value and two abstract arrays as arguments for
%dynamic analysis.  Then, dynamic analysis is successfully performed in line 2-8
%because it just checks the length of \jscode{arguments} and passes its elements
%to \jscode{array} and \jscode{args}.  Moreover, \jscode{isArray(array)} is also
%able to be evaluated using concrete semantics because we know that \code{array}
%is an abstract array.  Only copying via \jscode{copyArray}, flattening via
%\jscode{baseFlatten}, and pushing via \jscode{arrayPush} utilize the abstract
%semantics.  This is the core idea of lazy concrete execution to maximize the
%part of dynamic analysis during static analysis.

In the remaining section we formally define the combined analysis with sealed
symbolic execution in Section~\ref{sec:formal}.  Then, we explain how to define
the combined analysis for JavaScript programs in Section~\ref{sec:javascript}.