### 2020-11-02 paper/pldi21/abstract.tex
Revised intro

\begin{abstract}
  JavaScript becomes one of the most widely used programming languages for web
  development, server-side programming, and even micro-controllers for IoT.
  However, its dynamic language features degrade the scalability and precision
  of the static analysis.  Moreover, the variety of built-in functions and host
  environments requires excessive manual modeling of their behaviors.  To
  alleviate this problem, several researchers utilize dynamic analysis to take
  its advantage during JavaScript static analysis.  However, most of them cannot
  fully utilize high performance of dynamic analysis and sacrifice the soundness
  of static analysis.

  In this paper, we propose a novel \textit{dynamic shortcut} technique to
  leverage high performance of dynamic analysis for static analysis by freely
  switching between them in a sound way.  It could significantly accelerate the
  JavaScript static analysis by using the highly-optimized commercial JavaScript
  engines for the concretely executable program parts.  Moreover, it might
  improve the precision and lessen the modeling effort by performing dynamic
  analysis for opaque codes.  To maximize the usage of dynamic analysis, we
  propose \textit{sealed symbolic execution}, which is an augmented concrete
  execution that represents abstract values as sealed symbolic values.  We
  formally define the static analysis with dynamic shortcut using sealed
  symbolic execution.  We actualize our method via $\tool$, which is an
  extension of the existing JavaScript static analyzer SAFE with dynamic
  shortcut.  For evaluation of $\tool$, we analyzed abstracted versions of
  \inred{77} official tests of \inred{Lodash 4} library.  Our tool is
  \inred{X.X}x faster than the original static analysis on average.  Moreover,
  it improves \inred{XX.XX}\% of analysis precision by using dynamic analysis
  instead of static analysis for \inred{XX.X} opaque functions on average.
\end{abstract}