### 2020-11-03 paper/pldi21/motivation.tex
Revised motivation

\section{Motivation}\label{sec:motivation}

\begin{figure}[t]
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \begin{lstlisting}[style=myJSstyle]
function concat() {
  var length = arguments.length;
  if (!length) return [];
  var array = arguments[0],
      args  = Array(length - 1),
      index = length;
  while (index--)
    args[index-1] = arguments[index];
  return arrayPush(isArray(array) ?
    copyArray(array) : [array],
    baseFlatten(args, 1));
}
    \end{lstlisting}
    \vspace*{-1em}
    \caption{Lodash's \jscode{concat} function.}
    \label{fig:concat}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\textwidth}
    \begin{lstlisting}[style=myJSstyle,firstnumber=13]
function changeCountry(G) {
  ...
  if (G.selectedVal === "US" && state) {
    // deterministic arguments of `concat`
    state.items = _.concat([["Other", "Other"]],
      WebinarBase.questions.state.items);
    state.selectedVal = _.head(_.head(C.items));
  }
}
    \end{lstlisting}
    \vspace*{-1em}
    \caption{Load the list of states of the United States.}
    \label{fig:changeCountry}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\textwidth}
    \begin{lstlisting}[style=myJSstyle,firstnumber=22]
function getData(e) {
  var option = ... // option of server connection
  post(option).then(function(e) {
    if (e.total_records && e.total_records > 0) {
      // non-deterministic arguments of `concat`
      this.pastEvents =
        _.concat(this.pastEvents, e.events);
      this.total = e.total_records;
    } else this.noPastData = !0
  })
}
    \end{lstlisting}
    \vspace*{-1em}
    \caption{Load more Zoom events from the server.}
    \label{fig:getData}
  \end{subfigure}
  \vspace*{-1em}
  \caption{Motivating Example: Excerpts from Lodash library and JavaScript codes
  in \code{zoom.us} site.}
  \label{fig:example}
  \vspace*{-1em}
\end{figure}

In this section, we will explain the motivation of the dynamic shortcut for
JavaScript static analysis with real-world examples described in
Figure~\ref{fig:example}.  We first describe their program behaviors and then
explain how to utilize dynamic analysis during static analysis of them.

For motivating examples, we excerpt the \jscode{concat} function in
Figure~\ref{fig:concat} from Lodash library~\cite{lodash} (v4.17.20), which is
the most popular npm package\footnote{https://www.npmjs.com/browse/depended}
and \inred{124,562} npm packages have dependency with it.  The \jscode{concat}
function creates a new array concatenating given arrays or values.  It first
checks the length of arguments in line 1-3. Then, it stores the first argument
to \jscode{array} in line 4 and copies the remaining arguments to \jscode{args} in
line 5-8.  In line 9, it checks whether \jscode{array} is an array object with
the built-in function \jscode{isArray}.  If so, it creates a new array by
copying the given array via or initializing with a single value in line 10.
Finally, it flattens \jscode{args} via \jscode{baseFlatten} and pushes the
result to the new array in line 11.

For applications of Lodash, we excerpt two functions \jscode{changeCountry} in
Figure~\ref{fig:changeCountry} and \jscode{getData} in Figure~\ref{fig:getData}
from the \code{zoom.us}~\cite{zoom} site.  The website \code{zoom.us} is
homepage of Zoom, which is a videotelephony software program developed by Zooom
Video Communications and it is ranked as \inred{16th} popular web site according
to Alexa\footnote{https://www.alexa.com/siteinfo/zoom.us} in November 2020.

\paragraph{Dynamic shortcut with concrete values.}
When the given arguments of a function are concrete values, we can perform
dynamic analysis instead of static analysis. For example, the
\jscode{changeCountry} function is invoked when a user selects another country
on the drop-down list in the registration page.  It calls the \jscode{concat}
function to update the drop-down list of states or provinces in 17-18.  However,
when the user selects ``United States of America'' (USA), two arguments are
pre-defined with deterministic values; the first one is an array literal
\jscode{[["Other", "Other"]]} and the second one is an array of pairs of
abbreviations and names of the states defined as follows:
\begin{lstlisting}[style=myJSstyle,numbers=none]
WebinarBase.questions.state.items =
  [["AL","Alabama"], ..., ["WY", "Wyoming"]]
\end{lstlisting}
Moreover, \jscode{this} value is also a concrete value, the Lodash top-level
object \jscode{\_}.  Thus, we could perform dynamic analysis by invoking the
\jscode{concat} function with \jscode{\_} as \jscode{this} value and above two
concrete values as arguments.  It increases performance of static analysis by
skipping the analysis of function call in line 17-18 and utilizing the result of
dynamic analysis.

\begin{figure}[t]
  \begin{subfigure}{0.23\textwidth}
    \[
      \begin{array}{|c|c|}\hline
        \text{Property} & \text{Value}\\\hline
        \top & \symb_\jscode{evt}\\\hline
        \jscode{"length"} & \symb_\jscode{int}\\\hline
      \end{array}
    \]
    \vspace*{-1em}
    \caption{\jscode{this.pastEvents}}
    \label{fig:pastEvents}
  \end{subfigure}
  \begin{subfigure}{0.23\textwidth}
    \[
      \begin{array}{|c|c|}\hline
        \text{Property} & \text{Value}\\\hline
        \jscode{0} & \symb_\jscode{evt}\\\hline
        \cdots & \cdots\\\hline
        \jscode{7} & \symb_\jscode{evt}\\\hline
        \jscode{"length"} & \jscode{8}\\\hline
      \end{array}
    \]
    \vspace*{-1em}
    \caption{\jscode{e.events}}
    \label{fig:events}
  \end{subfigure}
  \vspace*{-1em}
  \caption{Concrete objects for arguments with sealed symbolic values.}
  \label{fig:sealed}
  \vspace*{-1em}
\end{figure}

\paragraph{Dynamic shortcut with abstract values.}
Dynamic analysis is still applicable using \textit{sealed symbolic execution}
even if the arguments are not concrete values.  The \jscode{getData} function is
invoked when clicking the ``Load More'' button to load more Zoom events in
``Webinars \& Events'' page.  For each click, the \jscode{getData} sends a POST
request to the server and receives additional event information \jscode{e} in
line 24.  Then, eight events in \jscode{e.events} are appended to
\jscode{this.pastEvents} using the \jscode{concat} function in line 27-28.
However, the arguments of \jscode{concat} are not deterministic because 1) the
event list stored in \jscode{this.pastEvents} is continuously grown for each
load and 2) also each event stored in \jscode{e.events} are dependent on the
data given from the server.

To perform dynamic analysis with abstract values, we sealed the abstract values
of arguments with symbolic values as described in Figure~\ref{fig:sealed}.  It
contains two symbolic values $\symbevt$ and $\symbint$ that represent any event
objects and integer values, respectively.  Then, dynamic analysis is
successfully performed before copying \jscode{array} via \jscode{copyArray} in
line 10.  First, \jscode{length} stores \jscode{2} and passes the length check
in line 1-2. Then, \jscode{array} points to the same object of
\jscode{this.pastEvents} in line 4, \jscode{args} stores an array with a
single object stored in \jscode{e.events} in 5-8, and the \jscode{isArray}
function returns \jscode{true} for \jscode{array} in line 9.  However, it fails
to perform dynamic analysis for \jscode{copyArray} because the \jscode{length}
property of \jscode{array} is the symbolic value $\symbint$.  Thus, the dynamic
shortcut returns the analysis result and the static analyzer continues to
analyze the program from line 10.  Then, only copying via \jscode{copyArray},
flattening via \jscode{baseFlatten}, and pushing via \jscode{arrayPush} utilize
the abstract semantics.  This is how to utilize sealed symbolic execution to
maximize the part of dynamic analysis during static analysis.

\paragraph{Dynamic shortcut for opaque functions.}

The previous two examples also show that dynamic shortcut can improve the
analysis precision and lessen the effort of modeling the opaque functions.
In line 9, the \jscode{isArray} function is a JavaScript built-in library thus
it is written in a native language of the host environment.  Thus, we need to
manually model its behavior to statically analyze it.  Assume that we model the
\jscode{isArray} function to return the top boolean value that denotes both of
\jscode{true} and \jscode{false}.  If we perform static analysis with this
modeling, both of true branch \jscode{copyArray(array)} and false branch
\jscode{[array]} in line 10 are always analyzed while \jscode{[array]} is never
reachable in the motivating examples.  However, with the dynamic shortcut, the
analyzer utilizes the concrete semantics of \jscode{isArray}.  It returns more
precise result \jscode{true} instead of the top boolean value and it is not
necessary to model the \jscode{isArray} function for static analysis of
motivating examples.