### 2020-11-15 paper/pldi21/conclusion.tex
Conclusion

\section{Conclusion}\label{sec:conclusion}

We have presented a novel technique \textit{dynamic shortcut} for JavaScript
static analysis.  It could significantly accelerate the JavaScript static
analysis by freely leveraging high performance of dynamic analysis for the
concretely executable program parts.  To maximize the usage of dynamic analysis,
we augmented the concrete execution to \textit{sealed symbolic execution} with
symbolic values for abstract values required abstract semantics.  We formally
defined the static analysis with dynamic shortcut using sealed symbolic execution
and proved its soundness on the formalization.  We developed $\tool$ as a
prototype implementation of the dynamic shortcut by combining state-of-the-art
static and dynamic analyzers SAFE and Jalangi.  Our tool accelerates the speed
of static analysis \inred{X.X}x for original tests and \inred{X.X}x for
abstracted tests of Lodash 4 library.  Moreover, it improves \inred{XX.XX}\% of
analysis precision by using dynamic analysis instead of static analysis for
\inred{XX.X} opaque functions on average.