### 2020-11-18 paper/pldi21/report/report.tex
Write soundness proof

\documentclass[11pt]{article}
\usepackage[a4paper,left=20mm,right=20mm,top=30mm,bottom=25mm]{geometry}

\usepackage{mathtools}
\usepackage[T1]{fontenc}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathpartir}
\usepackage{verbatim}
\usepackage{hyperref}
\usepackage{xspace}
\usepackage{stmaryrd}
\usepackage{tikz}
\usepackage{dirtree}
\usepackage{multicol}
\usepackage{array}
\usepackage{galois}
\usepackage{amsthm}

\usetikzlibrary{arrows}
\tikzset{
  treenode/.style = {align=center, inner sep=0pt, text centered},
  arn_n/.style = {treenode, circle, black, draw=black, text width=1.5em},
}

\begin{document}

\input{macros}

\title{Dynamic Shortcut Soundness Proof}

\author{Dongjun Youn$^*$ \and Jihyeok Park$^*$ \and Sukyoung Ryu$^*$}

\date{$^*$KAIST, South Korea}
\maketitle

\input{proof}

\end{document}