### 2020-11-19 paper/pldi21/report/report.tex

\documentclass[sigplan,10pt]{acmart}

\settopmatter{printacmref=false}

\usepackage[T1]{fontenc}
\usepackage[ruled, vlined]{algorithm2e}
\usepackage{amsfonts}
\usepackage{amsmath,amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{array}
\usepackage{booktabs}
\usepackage{bussproofs}
\usepackage{caption}
\usepackage{colortbl}
\usepackage{dirtree}
\usepackage{float}
\usepackage{galois}
\usepackage{graphicx}
\usepackage{hhline}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{makecell}
\usepackage{mathpartir}
\usepackage{mathtools}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{stmaryrd}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{tikz}
\usepackage{verbatim}
\usepackage{xcolor}
\usepackage{xspace}

\input{../macros}

\begin{document}

\title{Soundness Proof of Dynamic Shortcut}

\author{Anonymous Author(s)}
\date{}
% \author{Dongjun Youn$^*$ \and Jihyeok Park$^*$ \and Sukyoung Ryu$^*$}
% \date{$^*$KAIST, South Korea}

\maketitle

\input{def}
\input{proof}

\end{document}