### 2020-11-19 paper/pldi21/report/def.tex
Modified definition

\section{Definition of Dynamic Shortcut}\label{sec:def}

In this section, we introduce the formal definition of dynamic shortcut based on
sealed symbolic execution in abstract interpretation.

\subsection{Abstract Interpretation}

We first define the traditional abstract interpretation.

\subsubsection{Concrete Semantics}

We define a program $\prog$ as a state transition system $(\stset, \trans,
\istset)$.  A program starts with an initial state in $\istset$ and the
transition relation $\trans \subseteq \stset \times \stset$ describes how states
are transformed to other states.  A \textit{collecting semantics} $\sem{\prog} =
\{ \st \in \stset \mid \ist \in \istset \wedge \ist \trans^* \st \}$ consists of
reachable states from initial states of the program $\prog$.  We could calculate
it using the \textit{transfer function} $\transfer: \dom \rightarrow \dom$ as
follows:
\[
  \sem{\prog} = \underset{n \rightarrow \infty}{\lim}{\transfer^n(\ielem)}\\
  \qquad
  \transfer(\elem) = \elem \join \step(\elem)\\
\]
The \textit{concrete domain} $\dom = \powerset{\stset}$ is a complete lattice
with $\cup$, $\cap$, and $\subseteq$ as its join($\join$), meet($\meet$), and
partial order($\order$) operators.  The element $\ielem$ denotes the initial
states $\istset$.  The \textit{one-step execution} $\step: \dom \rightarrow
\dom$ transforms states using the transition relation $\trans$: $\step(\elem) =
\{ \st' \mid \st \in \elem \wedge \st \trans \st' \}$.

\subsubsection{Abstract Interpretation}

The abstract interpretation over-approximates the transfer $\transfer$ to the
\textit{abstract transfer function} $\abstransfer: \absdom \rightarrow \absdom$
to get the \textit{abstract semantics} $\abssem{\prog}$ in finite iterations as
follows:
\[
    \abssem{\prog} = \underset{n \rightarrow
    \infty}{\lim}{(\abstransfer)^n(\iabselem)}\\
\]
We define a \textit{state abstraction} $\dom \galois{\alpha}{\gamma} \absdom$ as
a Galois connection between the concrete domain $\dom$ and an abstract domain
$\absdom$ with a \textit{concretization function} $\gamma$ and an
\textit{abstraction function} $\alpha$.  The initial abstract state $\iabselem
\in \absdom$ represents an abstraction of the initial state set; $\ielem
\subseteq \gamma(\iabselem)$.  The abstract transfer function $\abstransfer:
\absdom \rightarrow \absdom$ is defined as $\abstransfer(\abselem) = \abselem
\join \absstep(\abselem)$ with an \textit{abstract one-step execution}
$\absstep: \absdom \rightarrow \absdom$.  For the sound state abstraction, the
join operator and the abstract one-step execution should satisfy the following
conditions:
\begin{itemize}
  \item $\forall \abselem_0, \abselem_1 \in \absdom. \; \gamma(\abselem_0) \cup
    \gamma(\abselem_1) \subseteq \gamma(\abselem_0 \join \abselem_1)$
  \item $\forall \abselem \in \absdom. \; \absstep \circ \gamma(\abselem) \subseteq
    \gamma \circ \absstep(\abselem)$
\end{itemize}

\subsubsection{Analysis Sensitivity}

Abstract interpretation is often defined with \textit{analysis sensitivity} to
increase the precision of static analysis.  A sensitive abstract domain
$\sabsdom: \viewset \rightarrow \absdom$ is defined with a \textit{view
abstraction} $\viewmap: \viewset \rightarrow \dom$ that provides multiple points
of views for reachable states during static analysis.  It maps a finite number
of views $\viewset$ to sets of states $\dom$. Each view $\view \in \viewset$
represents a set of states $\viewmap(\view)$.
A \textit{sensitive state abstraction} $\dom
\galois{\alpha_\viewmap}{\sgamma} \sabsdom$ is a Galois connection between
the concrete domain $\dom$ and the sensitive abstract domain $\sabsdom$ with the
following concretization function:
\[
  \sgamma(\sabselem) = \{ \st \in \stset \mid \forall \view \in \viewset.
  \; \st \in \viewmap(\view) \Rightarrow \st \in \gamma \circ \sabselem(\view) \}
\]

With analysis sensitivities, the abstract one-step execution $\sabsstep:
\sabsdom \rightarrow \sabsdom$ is defined as follows:
\[
  \sabsstep(\sabselem) = \lambda \view \in \viewset. \; \underset{\view' \in
  \viewset}{\bigjoin}{\viewtrans{\view'}{\view} \circ \sabselem(\view')}
\]
where $\viewtrans{\view'}{\view}: \absdom \rightarrow \absdom$ is the abstract
semantics of a \textit{view transition} from a view $\view'$ to another view
$\view$.  It should satisfy the following condition for the soundness of the
analysis:
\[
  \forall \abselem \in \absdom. \; \step(\gamma(\abselem) \cap \viewmap(\view'))
  \cap \viewmap(\view) \subseteq \gamma \circ
  \viewtrans{\view'}{\view}(\abselem)
\]

\subsection{Dynamic Shortcut}

We define dynamic shortcut by combining the sensitive abstract interpretation
with sealed symbolic execution.

\subsubsection{Sealed Symbolic Execution}

We define \textit{sealed symbolic execution} by extending the transition
relation $\trans$ as a symbolic transition relation $\symbtrans$ on symbolic
states.  First, we augment concrete states $\stset$ to symbolic states
$\symbstset$ by extending values $\valset$ with \textit{sealed symbolic values}
$\symbset$.  We also define the symbolic transition relation $\symbtrans
\subseteq \symbstset \times \symbstset$. We use the notation $\symbtrans^k$
for $k$ repetition of $\symbtrans$, and  $\symbst \symbtrans \excst$ when
$\symbst$ does not have symbolic transitions with any other symbolic states.  We
also define the soundness of sealed symbolic execution as follows:
\begin{definition}{(Soundness)}
  The symbolic transition relation is \textit{sound} when the following
  condition is satisfied:
  \[
    \begin{array}{l}
      \forall \imap \in \imapset. \; \forall \symbst, \symbst' \in \symbstset.\\
      \symbst \symbtrans \symbst' \Rightarrow \instant{\symbst}{\imap} \trans
      \instant{\symbst'}{\imap}
    \end{array}
  \]
  where $\imapset: \symbset \rightarrow \valset$ is instantiation maps from
  symblic values to real values, and $\instant{\symbst}{\imap}$ denotes a state
  produced by replacing each symbolic value $\symb$ to the corresponding value
  $\imap(\symb)$ using the instantiation map $\imap \in \imapset$.
\end{definition}

Note that the main difference of sealed symbolic execution with the traditional
symbolic execution is that it only supports sealed symbolic values instead of
symbolic expressions and path constraints.

We also define an \textit{abstract instantiation map} $\absimap \in \absimapset:
\symbset \rightarrow \absvalset$ with abstract values $\absvalset$.  Its
concretization function $\imapgamma: \absimapset \rightarrow
\powerset{\imapset}$ is defined with the concretization function $\valgamma:
\absvalset \rightarrow \powerset{\valset}$ for values as follows:
\[
  \imapgamma(\absimap) = \{
    \imap \mid \forall \symb \in \symbset. \;
    \imap(\symb) \in \gamma \circ \absimap(\symb)
  \}
\]
We define the instantiation of a given sealed symbolic state $\symbst \in
\symbstset$ with an abstract instantiation map $\absimap \in \absimapset$ as
follows:
\[
  \instant{\symbst}{\absimap} = \{ \instant{\symbst}{\imap} \mid \imap \in
  \imapgamma(\absimap) \}
\]

\begin{definition}[Sealed Symbolic Domain]
  We define \textit{sealed symbolic domain} $\symbdom: \powerset{\absimapset
  \times \symbstset}$ with the following concretization function $\symbgamma:
  \symbdom \rightarrow \dom$ and sealed symbolic one-step execution $\symbstep:
  \symbdom \rightarrow \symbdom$ as follows:
  \begin{itemize}
    \item \(
      \symbgamma(\symbelem) = 
      \bigcup \{ \instant{\symbst}{\absimap} \mid (\absimap, \symbst) \in
      \symbelem\}
    \)
    \item \(
      \symbstep(\symbelem) = \{ (\absimap, \symbst') \mid (\absimap, \symbst)
      \in \symbelem \wedge \symbst \symbtrans \symbst' \}
    \)
  \end{itemize}
\end{definition}

\subsubsection{Combined Domain}

\begin{definition}[Combined Domain]
  We define a \textit{combined domain} as $\combdom = \sabsdom \times \symbdom$
  with the following concretization function and join operator:
  \begin{itemize}
    \item \(
      \combgamma((\sabselem, \symbelem)) = \sgamma(\sabselem) \cup
      \symbgamma(\symbelem)
    \)
    \item \(
      (\sabselem, \symbelem) \join ({\sabselem}', \symbelem') = (\sabselem \join
      {\sabselem}', \symbelem \cup \symbelem')
    \)
  \end{itemize}
\end{definition}

An analysis element $\aelem \in \aelemset = (\viewset \times \absdom) \uplus
(\absimapset \times \symbstset)$ is either a pair of a view and an abstract
state or a sealed symbolic state.  To freely convert a pair of a view and an
abstract state to its corresponding pair of an abstract instantiation map and a
sealed symbolic state, we define two domain converters: \begin{itemize} \item
$\asconverter: (\viewset \times \absdom) \hookrightarrow (\absimapset \times
\symbstset)$ \item $\saconverter: (\viewset \times \absdom) \leftarrow
(\absimapset \times \symbstset)$ \end{itemize} The converter $\asconverter$ is a
\textit{partial function} and it is applicable when it is possible to represent
a given abstract state in a specific view as a pair of abstract instantiation
map and a sealed symbolic state.  Two converters should convert given elements
without loss of information:
\[
  \forall \view \in \viewset, \abselem \in \absdom.
\]
\[
  \asconverter((\view, \abselem)) = (\absimap, \symbst)
  \Rightarrow \left\{
  \begin{array}{l}
    (\view, \abselem) = \saconverter((\absimap, \symbst))\\\
    \viewmap(\view) \cap \gamma(\abselem) = \instant{\symbst}{\absimap}\\
  \end{array}
  \right.
\]
where $\identity$ denotes the identity function.

We define a $\atriage$ function using two converters between analysis elements
as follows:
\begin{definition}[$\atriage$]
  The triage function $\atriage: \aelemset \rightarrow \aelemset$ for analysis
  elements is defined as
  \[
    \atriage(\aelem) = \left\{
      \begin{array}{ll}
        \asconverter(\aelem)
        & \text{if} \; \aelem = (\view, \abselem) \wedge \aelem \in
        \Dom(\asconverter)\\
        \saconverter(\aelem)
        & \text{if} \; \aelem = (\absimap, \symbst) \wedge \symbst \symbtrans \symbst'\\
        \aelem
        & \text{Otherwise}
      \end{array}
    \right.
  \]
\end{definition}
Using the $\atriage$ function for analysis elements, we defined the $\triage$
function for combined states:
\begin{definition}[$\triage$]
  The triage function $\triage: \combdom \rightarrow \combdom$ for combined
  states is defined as
  \[
    \triage((\sabselem, \symbelem)) = \left(
      \lambda \view. \bigjoin \{ \abselem \!\mid\! (\view, \abselem) \in E \},
      E \cap (\absimapset \times \symbstset)
    \right)
  \]
  where
  \[
    E = \dot{\atriage}(\{ (\view, \abselem) \mid \view \in \viewset \wedge
    \sabselem(\view) = \abselem \} \cup \symbelem)
  \]
\end{definition}
The dot notation $\dot{f}$ denotes the element-wise extended function of a
function $f$.

\begin{definition}[Combined One-Step Execution]
  We define the combined one-step execution $\combstep: \combdom \rightarrow
  \combdom$ is define as follows:
  \[
    \combstep(\combelem) = (
      \sabsstep(\sabselem),
      \symbstep(\symbelem)
    )
  \]
  where $(\sabselem, \symbelem) = \triage(\combelem)$.
\end{definition}

We could configure when the sealed symbolic execution or the abstract
interpretation is performed based on the definition of the converter
$\asconverter$.  For example, we could define the $\asconverter$ for analysis
elements related to only function entry, call, and exit points to perform
conversions in a function level.  For the soundness and the termination of the
the abstract interpretation with dynamic shortcuts, it should satisfy the
following condition:
\begin{theorem}\label{theorem:shortcut}
  The abstract interpretation with dynamic shortcut is sound and terminates in a
  finite time if its abstract semantics is sound and $\saconverter$ satisfies the
  following condition:
  \[
    \asconverter((\view, \abselem)) = (\absimap, \symbst) \Rightarrow
    \symbst \symbtrans^k \excst \wedge 1 < k \leq N
  \]
  where $N$ is a pre-defined maximum length of the sealed symbolic execution.
\end{theorem}

In the remaining section, we will prove this theorem using the Kleene's
fixed-point theorem with the finite height of abstract domains.