### 2020-11-20 paper/pldi21/related.tex
Revised Sec. 7

\section{Related Work}\label{sec:related}
Researchers have introduced various analysis technique to utilize dynamic
analysis for static analysis in three ways: combined analysis,
automatic modeling, and pruning analysis scope.

\paragraph{Combined Analysis}
The most related previous work is combined analysis that utilizes dynamic
analysis during Java static analysis introduced by \citet{concerto}.
They proved that their combined analysis is sound and showed that it could
significantly improve the precision and performance of Java static analysis by
evaluating their tool, \concerto.  However, their approach has several
limitations compared with dynamic shortcuts.  First, it syntactically
divides a given program to \textit{applications} parts for static
analysis and \textit{frameworks} parts for dynamic analysis.  Thus, it is impossible to
freely switch between static analysis and dynamic analysis.  It is
even impossible to perform both static and dynamic analysis of 
the same program part in different contexts.  In addition, while they
introduced \textit{mostly-concrete interpretation} similar to our
sealed symbolic execution, it supports only a special \textit{unknown}
value that represents any possible value.  Thus, it cannot preserve
the precision of complex abstract domains~\cite{revisit-recency,
  regex, weaklySPE} frequently used in JavaScript static analysis.
On the contrary, sealed symbolic execution automatically detects when
to switch to static analysis to use abstract semantics for abstract values.
Finally, \concerto preserves the soundness when a program satisfies
the \textit{state separation hypothesis}.  It assumes that the states
of application parts and framework parts are not interrogated
or manipulated by each other.  While the assumption may be reasonable
for static analysis of Java applications using external libraries, it
is not satisfied for JavaScript programs in general.  Unlike their
approach, our approach does not have any assumptions between static and
dynamic analysis parts.

\paragraph{Automatic Modeling}
For static analysis of JavaScript programs, modeling behaviors of built-in
libraries or host-dependent functions is necessary because they are opaque code.
Since manual modeling is error-prone, tedious, and labor-intensive,
researchers~\cite{safewapi, safets} have utilized type information to
automatically model their behaviors.  However, type information is too imprecise
to reflect complex semantics and is difficult to represent the side-effects.
To alleviate the problem, \citet{mimic} introduced a technique
to infer JavaScript code for opaque code using concrete execution.
They captured the effects of opaque code on user objects by collecting partial
execution traces and synthesized JavaScript code using the extracted behaviors.
They also leveraged ES6 \jscode{Proxy} objects to capture the effects on user objects.
Instead of synthesizing JavaScript code,
\citet{opaque-model} presented a \textit{Sample-Run-Abstract (SRA)} approach for
on-demand modeling focusing on the current abstract states during static analysis.
It first \textit{samples} concrete states in a well-distributed manner,
\textit{runs} each sampled state on a JavaScript engine, and \textit{abstracts} execution results.
However, all the previous work sacrifice the soundness of static analysis.
On the contrary, while dynamic shortcuts may not be applicable to some
invocations of opaque functions, it is sound if it is applicable.

\paragraph{Pruning Analysis Scope}
Another approach to utilize dynamic analysis for JavaScript static analysis is
to prune the scope of analysis.  \citet{determinacy} proposed dynamic
determinacy analysis, which infers determinacy facts like whether a
variable or an expression at a given location holds the same value in
all possible execution of a program via concrete execution.
They specialized target source code with determinacy
facts so that static analysis can get benefits from elimination of dynamic code
generation such as \jscode{eval} and constant property names.
\citet{blended} introduced \textit{blended taint analysis}, which specializes JavaScript dynamic
language features such as dynamic code generation or variadic function calls.
It first performs dynamic analysis to collect traces with concrete values used
in dynamic language features and restricts the semantics of features based on the
collected traces during static analysis.  \citet{battles, eha} utilize three different points to reduce
analysis scope: initial states, dynamically loaded files, and event handlers.
They dumped the initial states from a specific web browser to focus on
analysis of the behaviors of web applications running on the browser.
Then, they collected paths of dynamically loaded files via concrete
execution and utilized the path information in static analysis.
For event handlers, they intentionally analyzed partial execution flows using concrete user events.
They collected concrete states for each entry of an event handler
during dynamic analysis, merged them to a single abstract state, and analyzed
the event handler with the abstract state.
Unfortunately, all the above mentioned approaches do not preserve
soundness of static analysis unlike our approach using dynamic shortcuts.

% \subsection{Abstract Counting}
% 
% \begin{itemize}
%   \item Improving flow analyses via $\Gamma$CFA: Abstract garbage collection and
%     counting~\cite{abstract-gc-counting}
%   \item Revisiting recency abstraction for JavaScript: towards an intuitive,
%     compositional, and efficient heap abstraction~\cite{revisit-recency}
% \end{itemize}

% \subsection{Access Analysis}
% 
% \begin{itemize}
%   \item Access Analysis-Based Tight Localization of Abstract
%     Memories~\cite{func-local}
%   \item Design and implementation of sparse global analyses for C-like
%     languages~\cite{sparse}
% \end{itemize}