### 2020-11-20 paper/pldi21/ds.tex
Section 3

\subsection{Combined Domain}

In this part, we define a \textit{combined domain} of a given sensitive abstract
domain with the sealed symbolic domain and its one-step execution.
\begin{definition}[Combined Domain]
  A \textit{combined domain} is $\combdom = \sabsdom \times \symbdom$ and its
  concretization function $\combgamma: \combdom \rightarrow \dom$ and join
  operator defined as
  \begin{align}
    \combgamma((\sabselem, \symbelem)) &= \sgamma(\sabselem) \cup
      \symbgamma(\symbelem)\\
    (\sabselem, \symbelem) \join ({\sabselem}', \symbelem') &= (\sabselem \join
      {\sabselem}', \symbelem \cup \symbelem')
  \end{align}
\end{definition}

Before defining the one-step execution for the combined domain, we introduce
\textit{analysis elements} to easily configure different types of abstract
states in sensitive abstract domain and sealed symbolic domain.
\begin{definition}[Analysis Elements]\label{def:aelem}
  An \textit{analysis element} $\aelem \in \aelemset = (\viewset \times \absdom)
  \uplus (\absimapset \times \symbstset)$ is either 1) a pair of a view and an
  abstract state in sensitive abstract domain $\sabsdom$, or 2) a pair of an
  abstract instantiation map and a sealed symbolic state in sealed symbolic
  domain $\symbdom$.  Its concretization function $\aelemgamma:
  \aelemset \rightarrow \dom$ is defined as
  \[
    \aelemgamma(\aelem) = \left\{
      \begin{array}{ll}
        \viewmap(\view) \cap \gamma(\abselem) & \text{if} \; (\view, \abselem) = \aelem\\
        \instant{\symbst}{\absimap} & \text{if} \; (\absimap, \symbst) = \aelem\\
      \end{array}
    \right.
  \]
\end{definition}

\begin{figure*}[t]
  \centering
  \begin{subfigure}[t]{0.15\textwidth}
    \includegraphics[height=3.2cm]{img/listing}
    \caption{Notations}
    \label{fig:ds-example1}
  \end{subfigure}
  \quad
  \begin{subfigure}[t]{0.23\textwidth}
    \includegraphics[height=3.2cm]{img/path-1}
    \caption{$\varx = 0$}
    \label{fig:ds-example2}
  \end{subfigure}
  \begin{subfigure}[t]{0.28\textwidth}
    \includegraphics[height=3.2cm]{img/path-2}
    \caption{$\varx > 0$}
    \label{fig:ds-example3}
  \end{subfigure}
  \begin{subfigure}[t]{0.28\textwidth}
    \includegraphics[height=3.2cm]{img/path-3}
    \caption{$\varx \in \mathbb{N}$}
    \label{fig:ds-example4}
  \end{subfigure}
  \caption{Three abstract interpretation using combined domain for the running
  example with different initial values for $\varx$.}
  \label{fig:ds-examples}
\end{figure*}

Moreover, to freely convert types of analysis elements, we define two converters:
\begin{align}
  \asconverter & : (\viewset \times \absdom) \hookrightarrow
    (\absimapset \times \symbstset)\\
  \saconverter & : (\viewset \times
    \absdom) \leftarrow (\absimapset \times \symbstset)
\end{align}
While the converter $\saconverter$ is total, the other one $\asconverter$ is
\textit{partial}. Thus, it is possible to convert an analysis element
$(\view, \abselem)$ in sensitive abstract domain to another analysis element in
sealed symbolic domain only if the convert $\asconverter$ is defined: $(\view,
\abselem) \in \Dom(\asconverter)$.  In addition, they should convert given
elements without loss of information for all $\aelem \in \aelemset$:
\[
  \asconverter(\aelem) = \aelem' \Rightarrow \left\{
  \begin{array}{l}
    \aelem = \saconverter(\aelem')\\
    \aelemgamma(\aelem) = \aelemgamma(\aelem')\\
  \end{array}
  \right.
\]

Now, we define the \textit{combined one-step execution} $\combstep: \combdom
\rightarrow \combdom$ with two converters $\asconverter$ and $\saconverter$ and
it consists of two steps; 1) the \textit{triage} step prepares the execution of
analysis elements by converting their types via two converters according to
their status, and 2) the \textit{execution} step performs execution of each
analysis element using abstract one-step execution $\absstep$ in sensitive
abstract domain and sealed symbolic one-step execution $\symbstep$ in sealed
symbolic execution.
\begin{definition}[Combined One-Step Execution]
  A \textit{combined one-step execution} $\combstep: \combdom \rightarrow
  \combdom$ is define as
  \[
    \combstep(\combelem) = (\sabsstep(\sabselem), \symbstep(\symbelem))
  \]
  where $(\sabselem, \symbelem) = \triage(\combelem)$.
\end{definition}

The $\triage$ function decides whether a new sealed symbolic execution is
\textit{introduced} or an existing one is \textit{terminated}.  When an analysis
element $(\view, \abselem)$ in the sensitive abstract domain is defined in the
converter $\asconverter$, a new sealed symbolic execution is introduced by
converting the analysis element to the corresponding one $(\absimap, \symbst) =
\asconverter((\view, \abselem))$ in the sealed symbolic domain.  In the other
hand, assume that if there is no more other sealed symbolic states having
symbolic transitions with $\symbst$ for an analysis element $(\absimap,
\symbst)$ in the sealed symbolic domain; $\symbst \symbtrans \excst$.  Then, the
sealed symbolic execution for $(\absimap, \symbst)$ is terminated by 1)
converting it to another analysis element $(\view, \abselem) =
\saconverter((\absimap, \symbst))$ in the sensitive abstract domain and 2)
merging with the current abstract state stored in the view $\view$ with
$\abselem$.

To formally define the $\triage$ function, we first define a $\atriage$ function
for analysis elements using two converters and then define the $\triage$ function.
\begin{definition}[$\atriage$]\label{def:atriage}
  The function $\atriage: \aelemset \rightarrow \aelemset$ for analysis elements
  is defined as
  \[
    \atriage(\aelem) = \left\{
      \begin{array}{ll}
        \asconverter(\aelem)
        & \text{if} \; \aelem = (\view, \abselem) \wedge \aelem \in
        \Dom(\asconverter)\\
        \saconverter(\aelem)
        & \text{if} \; \aelem = (\absimap, \symbst) \wedge \symbst \symbtrans
        \bot\\
        \aelem
        & \text{Otherwise}
      \end{array}
    \right.
  \]
\end{definition}
\begin{definition}[$\triage$]\label{def:triage}
  The triage function $\triage: \combdom \rightarrow \combdom$ for combined
  states is defined as
  \[
    \triage((\sabselem, \symbelem)) = \left(
      \lambda \view. \bigjoin \{ \abselem \!\mid\! (\view, \abselem) \in E \},
      E \cap (\absimapset \times \symbstset)
    \right)
  \]
  where
  \[
    E = \dot{\atriage}(\{ (\view, \sabselem(\view)) \mid \view \in \viewset \} \cup \symbelem)
  \]
  (The dot notation $\dot{f}$ denotes the element-wise extended function of a
  function $f$.)
\end{definition}

\subsection{Examples}

For examples of the abstract interpretation with the combined domain,
Figure~\ref{fig:ds-examples} depicts the flow of analysis for the running
example in Figure~\ref{fig:running-example} with three different initial set of
values for the variable $\varx$.  In this example, we use the abstract domain
$\{ -, 0, + \}$ for integers stored in $\varx$ introduced in
Section~\ref{sec:ai}, and the \textit{flow sensitivity} which utilizes the
labels of states as their views introduced in Section~\ref{sec:sens}.  For
brevity, we use the concatenated form to represent the abstract values such as
$-0$ denotes the set $\{ -, 0 \}$.

Figure~\ref{fig:ds-examples}(a) explains notations used in each graph. A solid
box denotes an analysis element, which is a pair of label $\lab$ and abstract
state $\abselem$.  A pair wrapped with angle brackets denotes an analysis
element, which is a pair of abstract instantiation map $\absimap$ and sealed
symbolic state $\symbst$.  In fact, the sealed symbolic state part (right) of
each pair in graphs contains only the value of the variable of $\varx$ without
the current label.  For brevity, we represents its label by locating them in the
next to the corresponding label.  A solid line is a view transition
$\viewtrans{\lab}{\lab'}$ from a label $\lab$ to another one $\lab'$.  A dotted
line is a sealed symbolic transition $\symbtrans$.  Three solid lines with
circled labels denote two converters $\saconverter$, $\asconverter$ and the join
operator $\join$.

Figure~\ref{fig:ds-examples}(b) shows the analysis with the combined domain when
the initial value of $\varx$ is $0$.  First, in the triage step,
the converter $\asconverter$ converts the analysis element $(\lab_0, 0)$ to
another analysis element $\langle \varnothing, 0 \rangle$ with the label
$\lab_0$.  Moreover, it does not introduce any sealed symbolic values because
the value represents only a single value.  Until the end of the program, the
sealed symbolic execution from $\langle \varnothing, 0 \rangle$ is successfully
continues.  However, there is no more possible symbolic transition for the
symbolic state $\langle \varnothing, 0 \rangle$ with the label $\lab_4$.  Thus,
it is converted to $(\lab_4, 0)$ via the converter $\saconverter$.

Instead of a single value, assume that the initial value of $\varx$ is one of
any positive integers.  Figure~\ref{fig:ds-examples}(c) describes the analysis
flow for such case.  Then, the initial abstract value at the label $\lab_0$ is
$+$ and it is impossible to convert to any sealed symbolic values because the
next program statement requires the actual value stored in the variable $\varx$
for the branch condition $\varx \geq 0$.  Thus, it performs view transition
$\viewtrans{\lab_0}{\lab_1}$ from the label $\lab_0$ to another one $\lab_1$ for
the abstract value $+$ and the result is also $+$.  Now, the analysis element
$(\lab_1, +)$ could be converted to $\langle \symb \mapsto +, \symb \rangle$
with the label $\lab_1$.  This sealed symbolic execution step terminates in the
label $\lab_3$ because the next statement is $\varx = -\varx$ and the negation
operator requires the actual value of $\varx$.  Finally, it performs the view
transition and the result is $(\lab_4, -)$.

For the last case, we assume that all integers are possible for the initial
value of the variable $\varx$ as described in Figure~\ref{fig:ds-examples}(d).
While it reaches the false branch in the label $\lab_2$ unlike previous cases,
it also cannot perform dynamic shortcuts because the statement in the false
branch is $\varx = -\varx$.  At the label $\lab_3$, there is two analysis
elements: 1) $(\lab_3, +)$ introduced by view transition from the label $\lab_2$
with $-$, and 2) $\langle \symb \mapsto 0+, \symb \rangle$ with $\lab_2$
introduced by sealed symbolic execution started at $\lab_1$.  However, since it
is not possible to perform sealed symbolic execution for both elements, the
second one is converted to $(\lab, 0+)$ and merged with $+$ at $\lab_3$ via the
join operator $\join$.  Finally, the view transition
$\viewtrans{\lab_3}{\lab_4}$ from $\lab_3$ to $\lab_4$ is performed to the
merged abstract state $0+$ and the result is $-0$.

\subsection{Soundness and Termination}

The converter $\asconverter$ and the sealed symbol transition $\symbtrans$ are
keys to configure the introduction and termination of sealed symbolic
executions.  To ensure the \textit{soundness} and \textit{termination} of the
abstract interpretation defined with the combined domain of sensitive abstract
domain and sealed symbolic domain, the following conditions should be hold.

\begin{theorem}[Soundness and Termination]\label{theorem:shortcut}
  The abstract interpretation with dynamic shortcut is \textbf{sound} and
  \textbf{terminates} in a finite time if
  \begin{itemize}
    \item The abstract transfer $\abstransfer$ is sound
    \item The sensitive abstract domain $\sabsdom$ has finite height
    \item The sealed symbolic transition $\symbtrans$ is valid
    \item There exists $N < \infty$ such that
      \[
        \forall \aelem \in \aelemset. \; \asconverter(\aelem) = (\absimap,
        \symbst) \Rightarrow \symbst
        \symbtrans^k \excst \wedge 1 < k \leq N
      \]
  \end{itemize}
\end{theorem}

To prove the termination of the abstract interpretation with the combined
domain, the condition provides an upper bound of the number of sealed symbolic
states after the sensitive abstract states are reached the fixed point.  Thus,
the termination of the analysis is guaranteed.  For the proof of Soundness, two
properties (\ref{equ:sound-join}) for join operator $\join$ and
(\ref{equ:sound-step}) for combined one-step execution are required to be
proved.  The core idea of the proof is to prove the following lemmas for
$\triage$ function and utilize them.
\begin{lemma}\label{lemma:step-symb}
  Assume that the following condition is hold
  \[
    \forall (\absimap, \symbst) \in \symbelem. \; \exists \symbst' \in
    \symbstset. \; \text{s.t.} \; \symbst \symbtrans \symbst'
  \]
  then the following property holds:
  \[
    \step \circ \symbgamma(\symbelem) \subseteq
    \symbgamma \circ \symbstep(\symbelem)\\
  \]
\end{lemma}
\begin{lemma}\label{lemma:triage}
  For a given combined state $\combelem \in \combdom$, the $\triage$ function
  satisfies the following two properties:
  \begin{itemize}
    \item $\combgamma(\combelem) \subseteq \combgamma \circ \triage(\combelem)$
    \item $\forall (\absimap, \symbst) \in \symbelem. \; \exists \symbst' \in
      \symbstset.  \; \text{s.t.} \; \symbst \symbtrans \symbst'$
  \end{itemize}
  where $(\sabselem, \symbelem) = \triage(\combelem)$
\end{lemma}