### 2020-11-20 paper/pldi21/shortcut.tex

\section{Dynamic Shortcut}\label{sec:formal}

In this section, we formally define dynamic shortcuts during static analysis by
combining abstract interpretation with sealed symbolic execution.  We extend the
formalization of abstract interpretation of \citet{abs-interp-1977,
abs-interp-1992} and views-based analysis sensitivity of \citet{sens-toplas}.
For dynamic shortcuts, we define a sealed symbolic execution and extend it to a
sealed symbolic domain with abstract instantiation maps.  To combine the
sensitive abstract interpretation with the sealed symbolic execution, we define
the combined domain of sensitive abstract domain and sealed symbolic domain and
explain it with a simple example. Finally, We prove the soundness and
termination property of abstract interpretation using the combined domain.

\subsection{Concrete Semantics}

We define a program $\prog$ as a state transition system $(\stset, \trans,
\istset)$.  A program starts with an initial state in $\istset$ and the
transition relation $\trans \subseteq \stset \times \stset$ describes how states
are transformed to other states.  A \textit{collecting semantics} $\sem{\prog} =
\{ \st \in \stset \mid \ist \in \istset \wedge \ist \trans^* \st \}$ consists of
reachable states from initial states of the program $\prog$.  We could calculate
it using the \textit{transfer function} $\transfer: \dom \rightarrow \dom$ as
follows:
\[
  \sem{\prog} = \underset{n \rightarrow \infty}{\lim}{\transfer^n(\ielem)}\\
  \qquad
  \transfer(\elem) = \elem \join \step(\elem)\\
\]
The \textit{concrete domain} $\dom = \powerset{\stset}$ is a complete lattice
with $\cup$, $\cap$, and $\subseteq$ as its join($\join$), meet($\meet$), and
partial order($\order$) operators.  The set of states $\ielem$ denotes the
initial states $\istset$.  The \textit{one-step execution} $\step: \dom
\rightarrow \dom$ transforms states using the transition relation $\trans$:
$\step(\elem) = \{ \st' \mid \st \in \elem \wedge \st \trans \st' \}$.

\begin{figure}[H]
  \[
    \begin{array}{r@{~}l@{~}c@{~}l}
      \labdot{0} & \kwif \; (\; \varx \geq 0 \;) & \labdot{1} & \varx = \varx ;\\
                 & \kwelse & \labdot{2} & \varx = -\varx ;\\
      \labdot{3} & \varx = -\varx ; & \labdot{4} \\
    \end{array}
  \]
  \vspace*{-1em}
  \caption{The negation of the absolute value of $\varx$}
  \label{fig:running-example}
\end{figure}

For example, the code in Figure~\ref{fig:running-example} is a simple program
that calculates the negation of the absolute value of the variable $\varx$.
States are pairs of labels and integers stored in $\varx$: $\stset = \labset
\times \mathbb{N}$.  Assume that the initial states are $\istset = \{ (\lab_0,
-42) \}$ which means the program starts with the variable $\varx$ that stores
$-42$ at $\lab_0$.  Then, the program is executed with the following trace:
\[
  (\lab_0, -42) \trans (\lab_2, -42) \trans (\lab_3, 42) \trans (\lab_4, -42)
\]

\subsection{Abstract Interpretation}\label{sec:ai}

The abstract interpretation~\cite{abs-interp-1977, abs-interp-1992}
over-approximates the transfer $\transfer$ to the \textit{abstract transfer
function} $\abstransfer: \absdom \rightarrow \absdom$ to get the
\textit{abstract semantics} $\abssem{\prog}$ in finite iterations as follows:
\[
    \abssem{\prog} = \underset{n \rightarrow
    \infty}{\lim}{(\abstransfer)^n(\iabselem)}\\
\]
We define a \textit{state abstraction} $\dom \galois{\alpha}{\gamma} \absdom$ as
a Galois connection between the concrete domain $\dom$ and an abstract domain
$\absdom$ with a \textit{concretization function} $\gamma$ and an
\textit{abstraction function} $\alpha$.  The initial abstract state $\iabselem
\in \absdom$ represents an abstraction of the initial state set; $\ielem
\subseteq \gamma(\iabselem)$.  The abstract transfer function $\abstransfer:
\absdom \rightarrow \absdom$ is defined as $\abstransfer(\abselem) = \abselem
\join \absstep(\abselem)$ with an \textit{abstract one-step execution}
$\absstep: \absdom \rightarrow \absdom$.  For the sound state abstraction, the
join operator and the abstract one-step execution should satisfy the following
conditions:
\begin{align}
  \forall \abselem_0, \abselem_1 \in \absdom & . \; \gamma(\abselem_0) \cup
  \gamma(\abselem_1) \subseteq \gamma(\abselem_0 \join
  \abselem_1) \label{equ:sound-join}\\
  \forall \abselem \in \absdom & . \; \absstep \circ \gamma(\abselem) \subseteq
  \gamma \circ \absstep(\abselem) \label{equ:sound-step}
\end{align}

A simple example abstract domain is $\absdom_\pm = \powerset{\{ -, +, 0 \}}$ with
set operators as domain operators; $-$ denotes negative integers, $+$ positive
integers, and $0$ zero.  Assume that we use it for the code in
Figure~\ref{fig:running-example} and the initial abstract state is $\iabselem =
\{ - \}$. Then, the analysis result becomes $\{ -, + \}$ because $\varx$ can
have positive values by executing $\varx = -\varx$ but there is no way to get
the value $0$ in this program.

\subsection{Analysis Sensitivity}\label{sec:sens}

Abstract interpretation is often defined with \textit{analysis sensitivity} to
increase the precision of static analysis.  A sensitive abstract domain
$\sabsdom: \viewset \rightarrow \absdom$ is defined with a \textit{view
abstraction} $\viewmap: \viewset \rightarrow \dom$ that provides multiple points
of views for reachable states during static analysis.  It maps a finite number
of views $\viewset$ to sets of states $\dom$. Each view $\view \in \viewset$
represents a set of states $\viewmap(\view)$ and also a state is always included
in a unique view: $\forall \st \in \stset. \; \exists! \view \in
\viewset.$ s.t. $\st \in \viewmap(\view)$.  A \textit{sensitive state
abstraction} $\dom \galois{\alpha_\viewmap}{\gamma_\viewmap} \sabsdom$ is a
Galois connection between the concrete domain $\dom$ and the sensitive abstract
domain $\sabsdom$ with the following concretization function:
\[
  \sgamma(\sabselem) = \underset{\view \in \viewset}{\bigcup}
  {\viewmap(\view) \cap \gamma \circ \sabselem(\view)}
\]

With analysis sensitivities, the abstract one-step execution $\sabsstep:
\sabsdom \rightarrow \sabsdom$ is defined as follows:
\[
  \sabsstep(\sabselem) = \lambda \view \in \viewset. \; \underset{\view' \in
  \viewset}{\bigjoin}{\viewtrans{\view'}{\view} \circ \sabselem(\view')}
\]
where $\viewtrans{\view'}{\view}: \absdom \rightarrow \absdom$ is the abstract
semantics of a \textit{view transition} from a view $\view'$ to another view
$\view$.  It should satisfy the following condition for the soundness of the
analysis:
\[
  \forall \abselem \in \absdom. \; \step(\gamma(\abselem) \cap \viewmap(\view'))
  \cap \viewmap(\view) \subseteq \gamma \circ
  \viewtrans{\view'}{\view}(\abselem)
\]

One of the most widely-used analysis sensitivity is \textit{flow sensitivity}
defined with the flow-sensitive view abstraction $\fsviewmap: \labset
\rightarrow \dom$ where:
\[
  \forall \lab\in\labset. \; \fsviewmap(\lab) = \{ \st \mid \st = (\lab, \_) \}
\]
If we apply the flow sensitivity for the above example with the initial abstract
state $[ \lab_0 \mapsto \{ -, 0, + \} ]$, the analysis result becomes as follows:
\[
  \begin{array}{|c||c|c|c|c|c|}\hline
    \labset & \lab_0 & \lab_1 & \lab_2 & \lab_3 & \lab_4\\\hline
    \absdom_\pm & -, 0, + & 0, + & - & 0, + & -, 0\\\hline
  \end{array}
\]

\subsection{Sealed Symbolic Execution}

We define \textit{sealed symbolic execution} by extending the transition
relation $\trans$ as a symbolic transition relation $\symbtrans$ on symbolic
states.  First, we augment concrete states $\stset$ to symbolic states
$\symbstset$ by extending values $\valset$ with \textit{sealed symbolic values}
$\symbset$.  We also define the symbolic transition relation $\symbtrans
\subseteq \symbstset \times \symbstset$. We use the notation $\symbtrans^k$
for $k$ repetition of $\symbtrans$, and  $\symbst \symbtrans \excst$ when
$\symbst$ does not have symbolic transitions with any other sealed symbolic
states.  We define the validity of sealed symbolic execution as follows:
\begin{definition}[Validity]\label{def:valid-symbtrans}
  The symbolic transition relation is \textit{valid} when the following
  condition is satisfied for any sealed symbolic states $\symbst$ and
  $\symbst'$:
  \[
    \symbst \symbtrans \symbst' \Leftrightarrow
    \forall \imap \in \imapset. \;
    \{ \st' \mid \instant{\symbst}{\imap} \trans \st' \}
    = \{ \instant{\symbst'}{\imap} \}
  \]
  where $\imapset: \symbset \rightarrow \valset$ represent \textit{instantiation
  maps} from symblic values to concrete values, and $\instant{\symbst}{\imap}$
  denotes a state produced by replacing each symbolic value $\symb$ to the
  corresponding value $\imap(\symb)$ using the instantiation map $\imap \in
  \imapset$.
\end{definition}

Note that The main difference of sealed symbolic execution with the traditional
symbolic execution~\cite{symbolic} is that it only supports sealed symbolic
values instead of symbolic expressions and path constraints.  For example, the
following trace represents the traditional symbolic execution of the running
example in Figure~\ref{fig:running-example}:
{
\small
\[
  \begin{array}{r@{~}c@{~}c@{~}c@{~}r@{~}c@{~}r}
    &&(\lab_1, \symb)[\symb \!\geq\! 0]
    &\trans& (\lab_3, \phantom{-}\symb)[\symb \!\geq\! 0]
    &\trans& (\lab_4, -\symb)[\symb \!\geq\! 0]
    \vspace*{-0.5em}\\
    &\rutrans&
    \vspace*{-0.5em}\\
    (\lab_0, \symb)[\varnothing]
    \vspace*{-0.5em}\\
    &\rdtrans&
    \vspace*{-0.5em}\\
    &&(\lab_2, \symb)[\symb \!<\! 0]
    &\trans& (\lab_3, -\symb)[\symb \!<\! 0]
    &\trans& (\lab_4, \phantom{-}\symb)[\symb \!<\! 0]\\
  \end{array}
\]
}
It first assigns a symbolic value $\symb$ to the variable $\varx$ at $\lab_0$.
For the conditional branch, it is forked by creating two symbolic states with
different path conditions $\symb \geq 0$ and $\symb < 0$ for true and false
branches, respectively.  After executing statements $\varx = \varx$ and $\varx =
-\varx$, the variable $\varx$ stores symbolic expressions $\symb$ and $-\symb$
at $\lab_3$. Similarly, $\varx$  stores $-\symb$ and $\symb$ at $\lab_3$.
However, the sealed symbolic execution stops at $\lab_0$ as follows:
\[
  (\lab_0, 0) \; \symbtrans \; \excst
\]
because the branch requires the actual value of the symbolic value $\symb$.

To define an abstract domain consists of sealed symbolic states, we define
\textit{abstract instantiation maps} $\absimapset: \symbset \rightarrow
\absvalset$ from symbolic values to abstract values.  Its concretization
function $\imapgamma: \absimapset \rightarrow \powerset{\imapset}$ is defined
with the concretization function $\valgamma: \absvalset \rightarrow
\powerset{\valset}$ for values as follows:
\[
  \imapgamma(\absimap) = \{
    \imap \mid \forall \symb \in \symbset. \;
    \imap(\symb) \in \gamma \circ \absimap(\symb)
  \}
\]
The instantiation of a given sealed symbolic state $\symbst \in \symbstset$ with
an abstract instantiation map $\absimap \in \absimapset$ is defined as follows:
\[
  \instant{\symbst}{\absimap} = \{ \instant{\symbst}{\imap} \mid \imap \in
  \imapgamma(\absimap) \}
\]
Now, we define the \textit{sealed symbolic domain} as follows:

\begin{definition}[Sealed Symbolic Domain]\label{def:symbdom}
  A \textit{sealed symbolic domain} $\symbdom: \powerset{\absimapset \times
  \symbstset}$ is defined with the following concretization function
  $\symbgamma: \symbdom \rightarrow \dom$ and sealed symbolic one-step execution
  $\symbstep: \symbdom \rightarrow \symbdom$ such that
  \begin{align}
    \symbgamma(\symbelem) &=
    \bigcup \{ \instant{\symbst}{\absimap} \mid (\absimap, \symbst) \in
    \symbelem\}\\
    \symbstep(\symbelem) &= \{ (\absimap, \symbst') \mid (\absimap, \symbst)
    \in \symbelem \wedge \symbst \symbtrans \symbst' \}
  \end{align}
\end{definition}