### 2020-11-20 paper/pldi21/javascript.tex
Section 4

\section{Dynamic Shortcut for JavaScript}\label{sec:javascript}

In this section, we introduce the core language of JavaScript that supports
first-class functions, open objects, and first-class property names, and define
sealed symbolic execution of the core language for dynamic shortcut.

\subsection{Core Language of JavaScript}

\[
  \begin{array}{ll@{~}c@{~}l}
    \text{Programs} & \prog &::=& (\lab: \inst)^*\\

    \text{Labels} & \lab &\in& \labset\\

    \text{Instructions} & \inst &::=&
    \refer = \expr \mid
    \refer = \kwobj \mid
    \refer = \expr ( \expr ) \mid
    \kwret \; \expr \mid
    \kwif \; \expr \; \lab\\

    \text{References} & \refer &::=&
    x \mid
    \expr [ \expr ]\\

    \text{Expressions} & \expr &::=&
    \pval \mid
    \lambda x. \; \lab \mid
    \refer \mid
    \op(\expr^*)\\
  \end{array}
\]

A program $\prog$ is a sequence of labeled instructions. An instruction $\inst$
is an expression assignment, an object creation, a function call, a return
instruction, or a branch.  A reference $\refer$ is a variable or a property
access of an object.  An expression $\expr$ is a primitive, a lambda function, a
reference, or an operation between other expressions.

\[
  \begin{array}{lr@{~}c@{~}l@{~}c@{~}l}
    \text{States} & \st &\in& \stset &=& \labset \times \memset \times
    \ctxtset \times \eaddrset\\
    \text{Memories} & \mem &\in& \memset &=& \locset \finmap \valset\\
    \text{Contexts} & \ctxt &\in& \ctxtset &=& \eaddrset \finmap (\eaddrset
    \times \labset \times \locset)\\
    \text{Locations} & \loc &\in& \locset &=& (\eaddrset \times \varset) \uplus
    (\oaddrset \times \strset)\\
    \text{Values} & \val &\in& \valset &=& \pvalset \uplus \oaddrset \uplus
    \fvalset\\
    \text{Primitives} & \pval &\in& \pvalset &=& \strset \uplus \cdots\\
    \text{Addresses} & \addr &\in& \addrset &=& \eaddrset \uplus \oaddrset\\
    \text{Functions} & \fval{x}{\lab} &\in& \fvalset &=& \varset \times
    \labset\\
  \end{array}
\]

States $\stset$ consist of labels $\labset$, memories $\memset$, contexts
$\ctxtset$, and environment addresses $\eaddrset$.  A memory $\mem \in \memset$
is a finite mapping from locations to values.  A context $\ctxt \in \ctxtset$ is
a finite mapping from environment addresses to tuple of environment addresses,
return labels, and left-hand side locations.  A location $\loc \in \locset$ is a
variable or an object property; a variable location consists of an environment
address and its name, and an object property location consists of an object
address and a string value.  A value $\val \in \valset$ is a primitive, an
address, or a function value.  An address $\addr \in \addrset$ is an environment
address or an object address.  A function value $\fval{x}{\lab} \in \fvalset$
consists of a parameter and a body label.  In the core language, the closed
scoping is used for functions for brevity thus only parameters and local
variables are accessible in the function body.

We formulate the concrete semantics of the core language.  The transition
relation between concrete states is defined with the semantics of references and
expressions using two different forms \fbox{$\referrule{\st}{\refer}{\loc}$} and
\fbox{$\exprrule{\st}{\expr}{\val}$}, respectively.  The initial states are
$\istset = \{ (\ilab, \varnothing, \epsilon, \tladdr) \}$ where $\ilab$ denotes
the initial label, $\epsilon$ empty map, and $\tladdr$ the top-level environment
address.  The function $\labnext$ returns the next label of a given label in the
current program $\prog$.

\subsection{Abstract Semantics}

In abstract semantics of the core language, we use the flow sensitivity with a
flow sensitive view abstraction $\fsviewmap: \labset \rightarrow \dom$ that
discriminates states using their labels: $\forall \lab \in \labset. \;
\fsviewmap(\lab) = \{ \st \in \stset \mid \st = (\lab, \_, \_, \_) \}$. Thus, the
sensitive abstract domain is defined as $\sabsdom = \labset \rightarrow
\absdom$.  We define an abstract state $\abselem \in \absdom$ as a tuple of an
abstract memory, an abstract context, an abstract address, and set of singleton
abstract addresses as follows:

\[
  \begin{array}{lr@{~}c@{~}l@{~}c@{~}l}
    \text{Abstract States} & \abselem &\in& \absdom &=& \absmemset \times \absctxtset
    \times \absaddrset \times \abscountset\\
    \text{Abstract Memories} & \absmem &\in& \absmemset &=& \abslocset \finmap
    \absvalset\\
    \text{Abstract Contexts} & \absctxt &\in& \absctxtset &=& \absaddrset \finmap
    \powerset{\absaddrset \times \viewset \times \powerset{\abslocset}}\\
    \text{Abstract Locations} & \absloc &\in& \abslocset &=& (\absaddrset \times
    \varset) \uplus (\absaddrset \times \strset)\\
    \text{Abstract Values} & \absval &\in& \absvalset &=& \powerset{\pvalset
    \uplus \absaddrset \uplus \fvalset}\\
    \text{Abstract Addresses} & \absaddr &\in& \absaddrset &=& \labset\\
    \text{Abstract Counters} & \abscount &\in& \abscountset &=& \absaddrset
    \rightarrow \{ \abszero, \absone, \absmany \}\\
  \end{array}
\]

An abstract memory $\absmem \in \absmemset$ is a finite mapping from abstract
locations $\abslocset$ to abstract values $\absvalset$.  Abstract locations
$\abslocset$ are pairs of abstract addresses with variable names or string
values. Abstract addresses $\absaddrset$ are defined with the
\textit{allocation-site abstraction} that partitions concrete addresses
$\addrset$ based on their allocation sites $\labset$.  Abstract contexts
$\absctxtset$ are finite maps from abstract addresses to powersets of triples of
abstract addresses, views, and powerset of abstract locations.  For abstract
counting~\cite{abstract-gc-counting, revisit-recency} in static analysis, we
define abstract counters $\abscountset$ that maps from abstract addresses to
their abstract counts representing how many times each abstract address has been
allocated; $\abszero$ denotes that it has never been allocated, $\absone$ once,
and $\absmany$ more than or equals to twice.

We define the semantics of view transition for the core language.  For abstract
memories, we use the notation $\absmem[L \mapstos \absval]$ to represent the
update of multiple abstract locations in $L$ with the abstract value $\absval$.
It performs the strong update if the abstract address for an abstract location
$(\absaddr, \_) \in L$ is singleton: $\abscount(\absaddr) = \absone$.
Otherwise, it performs the weak update for the soundness.  Moreover, we define
the increment function $\inc: \abscountset \times \absaddrset \rightarrow
\abscountset$ of the abstract counter defined as follows:
\[
  \inc(\abscount)(\absaddr_0) = \lambda \absaddr \in \absaddrset. \; \left\{
    \begin{array}{ll}
      \absone & \text{if} \; \absaddr = \absaddr_0 \wedge
      \abscount(\absaddr_0) = \abszero\\
      \absmany & \text{if} \; \absaddr = \absaddr_0 \wedge
      \abscount(\absaddr_0) = \absone\\
      \abscount(\absaddr) & \text{otherwise}
    \end{array}
  \right.
\]

\subsection{Sealed Symbolic Execution}

We define the symbolic states with the extended concrete values and the abstract
counters for addresses as follows:
\[
  \begin{array}{r@{~}c@{~}l}
    \symbstset &=& \labset \times \memset \times \ctxtset \times \eaddrset
    \times \abscountset\\
    \ctxtset &=& \eaddrset \finmap ((\eaddrset \times \labset \times \locset)
    \uplus \symbset)\\
    \valset &=& \pvalset \uplus \oaddrset \uplus \fvalset \uplus \symbset\\
    \abscountset &=& \oaddrset \rightarrow \{ \abszero, \absone, \absmany \}\\
  \end{array}
\]
We define the converter $\asconverter: (\viewset \times \absdom) \rightarrow
(\absimapset \times \symbstset)$ from pairs of views and abstract states to the
corresponding pairs of abstract instantiation maps and sealed symbolic states.
The algorithm is to traverse the averall structure, to remain a single value,
but changes to each abstract value represents multiple values to a new symbol
and adds the mapping from the new symbol to the original abstract value.  In a
reverse way, the opposite converter $\saconverter: (\absimapset \times
\symbstset)  \rightarrow (\viewset \times \absdom)$ recovers the abstract values
based on the given abstract instantiation map.  The symbolic transition relation
$\symbtrans$ is an extension of the original concrete transition relation
$\trans$.  When the exact values of symbolic values are required, symbolic
states have the symbolic transition relation with the exception state $\excst$.
For example, we add the following rule for $\kwret$ statements:
\begin{mathpar}
  \inferrule
  {
    \prog(\lab) = \kwret \; \expr\\
    \exprrule{\st}{\expr}{\val}\\
    \ctxt(\addr) \in \symbset
  }
  {
    \st = (\lab, \mem, \ctxt, \addr)
    \trans \excst
  }
\end{mathpar}
We extend each rule of concrete semantics to support such behaviors of symbolic
values.