### 2020-11-21 paper/pldi21/conclusion.tex

\section{Conclusion}\label{sec:conclusion}
We presented a novel technique for JavaScript static analysis using \textit{dynamic shortcuts}.
It can significantly accelerate the JavaScript static
analysis by freely leveraging high performance of dynamic analysis for
concretely executable program parts.  To maximally use dynamic analysis during static analysis,
we proposed \textit{sealed symbolic execution}, which performs
concrete execution using sealed symbolic values for abstract values.
We formally defined static analysis using dynamic shortcuts in the
abstract interpretation framework and proved its soundness and termination.
We developed $\tool$ as a prototype implementation of the proposed approach
by extending a combination of the state-of-the-art static and dynamic
analyzers SAFE and Jalangi.  Our tool accelerates the speed
of static analysis \inred{19.96}x for original tests and \inred{6.30}x for
abstracted tests of Lodash 4 library.  Moreover, it detect \inred{6} more of
dead branches by using sealed symbolic execution instead of
manual modeling for \inred{12} opaque functions on average.