### 2021-05-13 paper/fse21/conclusion.tex

\section{Conclusion}\label{sec:conclusion}
We presented a novel technique for JavaScript static analysis using \textit{dynamic shortcuts}.
It can significantly accelerate static
analysis by freely leveraging high performance of dynamic analysis for
concretely executable program parts.  To maximize such benefits,
we proposed \textit{{\sealed} execution}, which performs
concrete execution using {\sealed} values for abstract values.
We formally defined static analysis using dynamic shortcuts in the
abstract interpretation framework and proved its soundness and termination.
We developed $\tool$ as a prototype implementation of the proposed approach
by extending a combination of the state-of-the-art static and dynamic
analyzers SAFE and Jalangi.  Our tool accelerates the speed
of static analysis 22.30$\x$ for original tests and 7.81$\x$ for
abstracted tests of Lodash 4 library.  Moreover, it reduces the number
of failed assertions by 12.31\% by using {\sealed} execution instead of
manual modeling for 22 opaque functions on average.